---
title: "toy_example"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{toy_example}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

This document quickly shows how to use this package. For further details, see TBD (citation).

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
# devtools::install_github()
# devtools::build_rmd('vignettes/toy_example.Rmd')
devtools::load_all()
library(tidyverse)
```

First, setup. We use the bimodal family.
```{r}
den.fam <- biM_family()

grid.size <- 1024
grid <- seq(den.fam$range[1], den.fam$range[2], length.out = grid.size)

n.train <- 50
train.size <- 200
n.test <- 250
test.size  <- 10
max.k <- 10

# generate data
set.seed(42)
dat <- genSimData(
  den.fam,
  n.train, train.size,
  n.test, test.size,
  grid.size
)
```

Next, use pre-smoothing to turn discrete observations into density functions.
```{r}
dat$train.obsv <- dat$train.sample
# get bandwidth
bw <- quantile(apply(dat$train.obsv, 1, bw.SJ), 0.5)
dat$train.sample <- preSmooth.kde(
  obsv = dat$train.obsv,
  grid = grid,
  kde.opt = list(
    bw = bw, kernel = 'g',
    from = den.fam$range[1],
    to = den.fam$range[2]
  )
)
dat$bw <- bw
```

Further, transform the density functions into Hilbert space via centered log transformation (centered log-ratio).
```{r}
tm.c <- normL2(rep(1, grid.size), grid = grid) ^ 2 #constant used for centering
ls.tm <- toHilbert(
  dat$train.sample, grid = grid, 
  transFun = function(f, grid) orthLog(f, grid, against = 1 / tm.c),
  eps = .Machine$double.eps^(1/2)
)
dat$train.curve <- ls.tm$mat.curve
dat$train.idx.drop <- ls.tm$idx.drop
dat$ref.pt <- ls.tm$ref.pt
rm(ls.tm)
dat$grid <- grid
```

Functional principal component analysis and the induced density family.
```{r}
fpca.res <- do.call(
  fdapace::FPCA,
  list(
    Ly = asplit(dat$train.curve, 1),
    Lt = replicate(n = nrow(dat$train.curve), expr = grid, simplify = FALSE),
    optns = list(
      error = TRUE,
      lean = TRUE,
      FVEthreshold = 1,
      methodSelectK = 'FVE',
      plot = FALSE,
      useBinnedData = 'OFF'
    )
  )
)
fpca.den.fam <- fpca2DenFam(fpca.res, control = list(num.k = max.k))
```

We can now estimate using the induced family `fpca.den.fam`.
```{r}
ls.fpca.esti <- fpcaEsti(
  mat.obsv = dat$test.sample,
  fpca.res = fpca.res,
  esti.method = c("FPCA_BLUP", "FPCA_MAP", "FPCA_MLE"),
  control = list(
    num.k = 'AIC', max.k = max.k,
    method = 'LBFGS', return.scale = 'parameter'
  )
)
```
This is the estimated parameters, one row for one sample.
```{r}
ls.fpca.esti$res %>% head
```
This is the methods and additional information.
```{r}
ls.fpca.esti$idx %>% head
```

We can also look at the density functions.
```{r, fig.width=8, fig.height=5}
mat.pdf <- par2pdf(
  fpca.den.fam, 
  fpca.den.fam$fill.par(ls.fpca.esti$res %>% `[<-`(is.na(.), 0)),
  grid = grid
)
mat.pdf.true <- par2pdf(den.fam, dat$test.par, grid = grid)
df.pdf <- ls.fpca.esti$idx %>% mutate(idx.row = row_number()) %>% 
  mutate(
    pdf = map(idx.row, ~ {
      tibble(x = grid, pdf = mat.pdf[.x, ])
    })
  ) %>% as_tibble
df.pdf <- df.pdf %>% bind_rows(
  tibble(idx.obsv = seq(n.test), esti.method = 'true') %>% 
    mutate(pdf = map(idx.obsv, ~ {
      tibble(x = grid, pdf = mat.pdf.true[.x, ])
    }))
)
df.pdf %>% 
  filter(idx.obsv %in% sample.int(n.test, 4)) %>% 
  unnest('pdf') %>% 
  mutate(esti.method = factor(esti.method, c(
    'true', 'FPCA_MLE', 'FPCA_MAP', 'FPCA_BLUP'
  ))) %>% 
  ggplot(aes(x, y = pdf)) +
  geom_path(aes(
    color = esti.method, linetype = esti.method
  )) +
  labs(x = 'x', y = 'density function', color = 'method', linetype = 'method') +
  facet_grid(~ idx.obsv, labeller = label_both) +
  theme(legend.position = 'bottom')
```

